---
title: "The Atlantic Hydra and how to subdue (not kill) it"
author: "Daniel Ricard - DFO Science Gulf Region"
date: "2022-12-12"
output: 
  beamer_presentation:
    theme: "Madrid"
    colortheme: "seahorse"
    fonttheme: "structurebold"
    slide_level: 2
urlcolor: blue
header-includes:
- \AtBeginDocument{\title[Atlantic Hydra]{The Atlantic Hydra and how to subdue it}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(gulf)
```

## the Atlantic Hydra


* also known as the Species List

* one can idealistically assume that an authoritative and curated list of species exists regionally

* the Species List is known to duplicate and diverge each time a new usage case is formulated

* because of changes in species distributions and in taxonomic research, the Species List is continually evolving

## Timeline of Atlantic Hydra sightings

* Appendix in [Hurlbut and Clay (1990)](https://cat.fsl-bsf.scitech.gc.ca/record=b3882665~S1) states:

This file is modified from one provided in 1987 by the Marine Fish Division, St. Andrews, N.B.

* R package *gulf*
  - data frame called *species.names*, and helper functions *species.str* and *species.code*
  - assignment of AphiaID to existing list, "beautification" of common names, ...

* [Maritimes Atlas 2022](https://cat.fsl-bsf.scitech.gc.ca/record=b4106600~S1)

* laboratory processing of stomach contents: the RV species list as a starting point?

* Andes deployment in 2022 annual ecosystem survey following merge of species lists between Quebec and Gulf

## Attempts at subduing the Hydra

* an adhoc but reproducible solution : Maritimes Atlas
  - the summary of catch observations from the summer survey is the starting point, but there are a number of *ad hoc* interventions done to [generate Table 1](https://github.com/dfo-gulf-science/Maritimes-SUMMER-Atlas/blob/main/summaries-catch-records.R)

* "select few" solution : SAS to R report, use a list of species that was used previously in a SAS script

* a flexible solution in the gulf package, using functions *taxonomic.group* and *years.recorded*, to obtain a defensible list of species and years when indices of abundance and biomass, and distribution maps can be generated

* this workshop

## WoRMS to extract taxonomic details from an AphiaID

```{r}
t.df <- worms::wormsbyid(
  species.names[which(species.names$code==10),"aphia_id"]
)
print(t.df)
```


* knowing the taxonomic rank of an observation allows us to establish a hierarchy

* the taxonomic hierarchy opens the door to grouping observations based on a lower taxonomic rank


## 1 - all possible combinations of years and species

for all possible species $s$ and year $y$, define an $s \times y$ matrix that denotes all possible combinations, and set all of them equal to zero

\begin{equation}
S_{all}
\end{equation}


## 2 - catch records

using the database catch records, generate another $s$ and year $y$ matrix showing what species codes are used in the database catch records

\begin{equation}
S_{observed}
\end{equation}

* this matrix should have smaller dimensions than $S_{all}$
* values of zero denote that no observations of a given species was made in a given year

## 3 - time period with consistency of observations

* for each species code that we see in the database catch table, note the years when it was/is expected to be recorded

* examples for shrimps, sponges and small cod

\begin{equation}
S_{recorded}
\end{equation}


## 4 - aggregation by groups of species

define an $s$ by $s$ matrix that encodes the species groupings

* add a third dimension, the taxonomic rank

* lower the rank of observations and add the catches for the group

* this is what is done in our "SAS to R" report

* example with *Liparis* species
* example with wolffish species

## summary of Gulf Region 2022 post-survey fixes required

- before the survey, creation of new 5-digit codes for cases where no merge could be achieved between Quebec and Gulf

* few fixes required for fish species, but it was mainly for invertebrates
  - non-taxonomic 
  - taxonomic precision
  - discrepancy between ID guide and Andes


## identification guides

* what we mean we say "the invertebrate identification guide": 
[Nozeres, C. 2015. St. Lawrence marine invertebrates posters. figshare.](https://dx.doi.org/10.6084/m9.figshare.985581.v5)

Nozères, C. et Archambault, D. 2014. Portfolio pour l'identification rapide
d'invertébrés capturés au chalut dans l'estuaire et le nord du golfe du Saint-Laurent.
Rapp. manus. can. sci. halieut. aquat. 3033 : iv + 30 p.


* what we mean we say ["the sponge identification guide"](https://publications.gc.ca/site/eng/9.888681/publication.html)


## What can we learn from Heracles' experience?

* Heracles was helped by his nephew Iolaus

* if killing the Hydra is the goal, burn its neck using a sword and fire (this seemingly stops the regrowth of heads)

* since killing the Hydra is not the goal here, let's articulate an alternative course of action

\begin{center}
  \includegraphics[width=1.0\textwidth]{M13.1Hydra_sm.png}
\end{center}

## For discussion

* what is the citable source for the list used in the Gulf and Maritimes RV surveys?

* worst case scenario: each user had its own copy of a distant list and tweaks it to suit its own needs

* best case scenario: 


## Discussion

\begin{center}
  \includegraphics[width=1.0\textwidth]{teleost-navire-stock-poissons-jacques-cartier.png}
\end{center}
